A Trio of 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia

### 2022-04-05 papers/ms.tex
update ms
mostly tables
some text

\title{
  Three 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
}

@@

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new and
  refined view of open cluster dispersal.
  Here we present an analysis of a group of $\approx$40 million
  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.
  The group includes four known Kepler Objects of Interest:
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
  Kepler-1627 is a Neptune-sized planet in a sub-component of the
  complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.
  Here we focus on the latter three systems, which are in other
  sub-components of the complex (RSG-5 and CH-2).
  Based on kinematic evidence from Gaia, stellar rotation periods from
  TESS, and spectroscopy, these three systems are also $\approx$40
  million years old.
  More specifically, we find that Kepler-1643 (in RSG-5) is
  $46^{+9}_{-7}$\,Myr old, while KOI-7368 and KOI-7913 (in CH-2) are
  $36^{+10}_{-8}$\,Myr old.
  Based on the transit shapes and high resolution imaging, they are
  all most likely planets, with false positive probabilities of
  $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
  Kepler-1643, KOI-7368, and KOI-7913 respectively.
  These planets therefore empirically
  demonstrate that mini-Neptunes with sizes of $\approx$2 Earth
  radii exist at ages of $\approx$40 million years.
\end{abstract}

@@

In this work, we expand on our previous study of a $38^{+7}_{-6}$
million year old Neptune-sized planet in the Kepler field (Kepler-1627
Ab; \citealt{bouma_kep1627_2022}).  The age of this planet was derived
based on its host star's membership in the $\delta$\ Lyr\ cluster.
Our analysis of the cluster focused on the immediate spatial and
kinematic vicinity of Kepler-1627~A in order to confirm the age of the
planet.  However it became clear that the $\delta$\ Lyr\ cluster seems
to also be part of a much larger group of similarly aged stars.  This
group, which is at a distance of $\sim$330 parsecs from the Sun,
appears to span Cepheus to Hercules (galactic longitudes, $l$, between
40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
0$^\circ$ and 20$^\circ$.  We therefore refer to it as the Cep-Her
complex.  It exhibits significant sub-structure over its $\approx$250
parsec length, and a detailed analysis of its memberships, kinematics,
and possible origin is currently being prepared by R.~Kerr and
collaborators.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Positions and velocities of candidate members of the Cep-Her
  complex.}
  {\it Top row}: On-sky positions in galactic coordinates.  Black
  points are stars for which group membership is more secure than for
  gray points (see Section~\ref{subsec:members}).
  Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster
  \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\} \approx
  \{ 66^\circ, 12^\circ\}$.
  {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
  \{0, 0, 20.8\}$\,pc; lines of constant heliocentric distance are
  shown between 250 and 400\,pc, spaced by 50\,pc.
  {\it Bottom row}: Galactic tangential velocities (left) and
  longitudinal tangential velocity versus galactic latitude (right).
  The gray band in the lower-right shows the $\pm$1-$\sigma$
  projection of the Solar velocity with respect to the local standard
  of rest.  There is a strong spatial and kinematic overlap between
  Kepler-1643 and RSG-5 ($\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$;
  $\{ X, Y, Z\} \approx \{60, 330, 50\} $ pc).  The local population
  of candidate young stars around KOI-7368 and KOI-7913 is more
  diffuse -- we call this region ``CH-2''.
  The {\bf interactive figure} enables a few different cuts to be
  shown.
	\label{fig:XYZvtang}
	}
\end{figure*}

@@

Our focus is on a region of the Galaxy approximately 200 to 500
parsecs from the Sun, above the galactic plane, and spanning galactic
longitudes of roughly $40^\circ$ to $100^\circ$ degrees.  Two rich
clusters in this region are the $\delta$~Lyra cluster
\citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
Each of these clusters was known before Gaia.  They have reported ages
between $\approx$30 and $\approx$60 million years.  Early empirical
evidence that these two clusters could be part of a large and more
diffuse population was apparent in the Gaia-based photometric analysis
of pre-main-sequence stars by \citet[][see their Figures~11
and~13]{Zari2018}.  Further kinematic connections and complexity were
highlighted by \citet{KounkelCovey2019}, who included these previously
known groups in the larger structures dubbed ``Theia~73'' and
``Theia~96''\footnote{See their visualization online at
\url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March 2022)}.
The connection made by \citet{KounkelCovey2019} between the previously
known open clusters and the other groups in the region was made as
part of an unsupervised clustering analysis of the Gaia DR2 positions
and tangential velocities with a subsequent manual ``stitching'' step,
and generally supports the idea that there is an overdensity of
$\approx$30 and $\approx$60 million year old stars in this region of
the Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the
Gaia DR2 point-source catalog out to one third of a kiloparsec,
identified three of the nearest sub-populations, dubbed
``Cepheus-Cygnus'', ``Lyra'', and ``Cerberus''.  \citet{Kerr2021}
reported ages for each of these subgroups between 30 and 35 million
years.

@@

The second step is to perform in an unsupervised HDBScan
clustering on the photometrically selected population
\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
parameters we use in this clustering analysis are $\{ X, Y, Z, c v_b,
c v_{l^*} \} $, where $c$ is the size-velocity corrective factor,
which is taken as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the
spatial and velocity scales have identical standard deviations.
Positions are computed assuming the \texttt{astropy v4.0} coordinate
standard \citep{astropy_2018}, which places the Sun $8122$ pc from the
galactic center, and assumes the solar velocity with respect to the
local standard of rest from \citet{schonrich_local_2010}.  As input
parameters to HDBScan, we set the minimum $\epsilon$ threshold past
which clusters cannot be fragmented as $25$ parsecs in physical space,
and $c$ \kms\ in velocity.  The minimum cluster size $N$ is set to 10,
as is $k$, the parameter used to define the ``core distance'' density
metric.

@@

Figure~\ref{fig:XYZvtang} shows the positions of the KOIs along
various projections.  Kepler-1643 is near the core RSG-5 population
both spatially and kinematically.  KOI-7368 and KOI-7913 are in a 
diffuse region $\approx$40 parsecs above RSG-5 in $Z$ and $\approx$100
parsecs closer to the Sun in $Y$.  In tangential galactic velocity
space, there may be some kinematic overlap between the region the
latter two KOIs are in, and the main RSG-5 group.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		%    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
		%    markers = ['P', 'v', 'X', 'X', 's']
		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
		Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
		its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
		KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of the
		youngest transiting planets could be explained by their primordial
		atmospheres not yet having evaporated; direct measurements of the
		atmospheric outflows or planetary masses would help to confirm
		this expectation.  Selection effects may also be important.
		Parameters are from the NASA Exoplanet Archive (2022 Feb 27).
		\label{fig:rp_period_age}
	}
\end{figure*}

@@

Perhaps the most significant point about
Kepler-1643~b, KOI-7368.01, and KOI-7913 is that they have normal
sizes.
{\bf TODO for LGB: write more about this}

@@

We have argued that Kepler-1643~b, KOI-7368.01, and KOI-7913 {\it i)}
are all $\approx$40 million years old, and {\it ii)} that each signal
is most likely planetary.  The evidence for the planetary
interpretation comes from an application of \texttt{VESPA} to the
Kepler data, alongside new adaptive optics imaging from NIRC2.  The
validity of the \texttt{VESPA} framework rests on the premise that
non-astrophysical false positives, including false alarms and known
neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
KOI-7368 both pass all but one of the canonical vetting checks: the
check for centroid offsets between the in and out-of-transit phases.
For both cases, we have argued the weak observed shifts are most
likely caused by starspot-induced variability in specific quarters
shifting the stellar center-of-light.

Each system has multiple indicators of youth that support their
reported ages.  For Kepler-1643, perhaps the strongest youth indicator
is its physical and kinematic association with RSG-5.  Based on the
color-absolute magnitude diagram, we seem to be able to select members
of this cluster with a false positive rate of $\ll 5\%$
(Figure~\ref{fig:age}), and Kepler-1643 is one such member.  Its
rotation period period agrees with this assessment, and {\bf its
lithium is a little low}.

The spatial and kinematic arguments for the youth of KOI-7368 and
KOI-7913 are necessarily weaker because they are in an association of
stars, CH-2, that is much more tenuous.  For KOI-7913 this is not of
particular concern, since it happens to be a binary.  Both components
have isochronal ages consistent with $40$\,Myr, but the measurement
for the secondary is more constraining since it is less massive.  Both
components also show H$\alpha$ in emission, which for the $\approx$K6V
primary is a good indicator that the star is $\lesssim$100$\,$Myr old.
For KOI-7368, which is slightly more massive, the Li 6708\AA\ EW
measurement and stellar rotation period are more constraining, and
provide independent verification of the stellar youth.

>>>>

\title{
  A Trio of 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
}

@@

% 208 words (250 max)
\begin{abstract}
  Stellar positions and velocities from Gaia are yielding a new and
  refined view of open cluster dispersal.
  Here we present an analysis of a group 
  stars spanning Cepheus ($l=100^\circ$) to Hercules
  ($l=40^\circ$), hereafter the Cep-Her complex.
  The group includes four Kepler Objects of Interest:
  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
  The latter Neptune-sized planet is in a sub-component of the
  Cep-Her complex called the $\delta$\ Lyr\ cluster
  \citep{bouma_kep1627_2022}.
  Here we focus on the former three systems, which are in other
  sub-components of the complex.
  Based on kinematic evidence from Gaia, stellar rotation periods from
  TESS, and spectroscopy, these three systems are also $\approx$40
  million years old.
  More specifically, we find that Kepler-1643 is in a dense
  sub-cluster of the complex called RSG-5, and is $46^{+9}_{-7}$\,Myr old.
  KOI-7368 and KOI-7913 are in a diffuse region that we call
  CH-2, and are $36^{+10}_{-8}$\,Myr old.
  Based on the transit shapes and high resolution imaging, all three are
  most likely planets, with false positive probabilities of
  $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
  Kepler-1643, KOI-7368, and KOI-7913 respectively.
  These planets therefore empirically
  demonstrate that mini-Neptunes with sizes of $\approx$2 Earth
  radii exist at ages of $\approx$40 million years.
\end{abstract}

@@

In this work, we expand on our previous study of a $38^{+7}_{-6}$
million year old Neptune-sized planet in the Kepler field (Kepler-1627
Ab; \citealt{bouma_kep1627_2022}).  The age of this planet was derived
based on its host star's membership in the $\delta$\ Lyr\ cluster.
Our analysis of the cluster focused on the immediate spatial and
kinematic vicinity of Kepler-1627~A in order to confirm the age of the
planet.  However it became clear that the $\delta$\ Lyr\ cluster seems
to also be part of a much larger group of similarly aged stars.  This
group, which is at a distance of $\sim$330\,pc from the Sun,
appears to span Cepheus to Hercules (galactic longitudes, $l$, between
40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
0$^\circ$ and 20$^\circ$.  We therefore refer to it as the Cep-Her
complex.  It exhibits significant sub-structure over its $\approx$250
parsec length, and a detailed analysis of its memberships, kinematics,
and possible origin is currently being prepared by R.~Kerr and
collaborators.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Positions and velocities of candidate members of the Cep-Her
  complex.}
  {\it Top row}: On-sky positions in galactic coordinates.  Black
  points are stars for which group membership is more secure than for
  gray points.
  Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster
  \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\} \approx
  \{ 66^\circ, 12^\circ\}$.
  {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
  \{0, 0, 20.8\}$\,pc; lines of constant heliocentric distance are
  shown between 250 and 400\,pc, spaced by 50\,pc.
  {\it Bottom row}: Galactic tangential velocities (left) and
  galactic longitudinal velocity versus galactic longitude (right).
  The gray band in the lower-right shows the $\pm$1-$\sigma$
  projection of the Solar velocity with respect to the local standard
  of rest.  There is a strong spatial and kinematic overlap between
  Kepler-1643 and RSG-5 (magenta).  The local population
  of candidate young stars around KOI-7368 and KOI-7913 is more
  diffuse -- we call this region ``CH-2'' (lime-green).
  The selection method for the black, gray, purple, and green points
  is described in Section~\ref{subsec:members}.
  %The {\bf interactive figure} enables a few different cuts to be
  %shown.
	\label{fig:XYZvtang}
	}
\end{figure*}

@@

Our focus is on a region of the Galaxy approximately 200 to 500\,pc
from the Sun, above the galactic plane, and spanning galactic
longitudes of roughly $40^\circ$ to $100^\circ$ degrees.  Two rich
clusters in this region are the $\delta$~Lyra cluster
\citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
Each of these clusters was known before Gaia.  They have reported ages
between $\approx$30 and $\approx$60 million years.  Early empirical
evidence that these two clusters could be part of a large and more
diffuse population was apparent in the Gaia-based photometric analysis
of pre-main-sequence stars by \citet[][see their Figures~11
and~13]{Zari2018}.  Further kinematic connections and complexity were
highlighted by \citet{KounkelCovey2019}, who included these previously
known groups in the larger structures dubbed ``Theia~73'' and
``Theia~96''\footnote{See their visualization online at
\url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March 2022)}.
The connection made by \citet{KounkelCovey2019} between the previously
known open clusters and the other groups in the region was made as
part of an unsupervised clustering analysis of the Gaia DR2 positions
and tangential velocities with a subsequent manual ``stitching'' step,
and generally supports the idea that there is an overdensity of
$\approx$30 and $\approx$60 million year old stars in this region of
the Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the
Gaia DR2 point-source catalog out to one third of a kiloparsec,
identified three of the nearest sub-populations, dubbed
``Cepheus-Cygnus'', ``Lyra'', and ``Cerberus''.  \citet{Kerr2021}
reported ages for each of these subgroups between 30 and 35 million
years.

@@

The second step is to perform an unsupervised HDBScan
clustering on the photometrically selected population
\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
parameters we use in this clustering analysis are $\{ X, Y, Z, c v_b,
c v_{l^*} \} $, where $c$ is the size-velocity corrective factor,
which is taken as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the
spatial and velocity scales have identical standard deviations.
Positions are computed assuming the \texttt{astropy v4.0} coordinate
standard \citep{astropy_2018}, which places the Sun $8122$ pc from the
galactic center, and assumes the solar velocity with respect to the
local standard of rest from \citet{schonrich_local_2010}.  As input
parameters to HDBScan, we set the minimum $\epsilon$ threshold past
which clusters cannot be fragmented as $25$\,pc in physical space,
and $c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10,
as is $k$, the parameter used to define the ``core distance'' density
metric.

@@

Figure~\ref{fig:XYZvtang} shows the positions of the KOIs along
various projections.  Kepler-1643 is near the core RSG-5 population
both spatially and kinematically.  KOI-7368 and KOI-7913 are in a 
diffuse region $\approx$40\,pc above RSG-5 in $Z$ and $\approx$100\,pc
closer to the Sun in $Y$.  In tangential galactic velocity space,
there may be some kinematic overlap between the region the
latter two KOIs are in, and the main RSG-5 group.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		%    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
		%    markers = ['P', 'v', 'X', 'X', 's']
		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
    3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
    its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
    KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of the
    youngest transiting planets could be tied to a selection effect.
    Kepler-1643, KOI-7368, and KOI-7913 have normal mini-Neptune sizes
    of 2 to 3\,$R_\oplus$, which is a novelty given their ages.
    Parameters are from the NASA Exoplanet Archive (2022 Apr 5).
		\label{fig:rp_period_age}
	}
\end{figure*}

@@

In the context of the existing exoplanet population, the most significant
novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is that they have normal
mini-Neptune sizes.
More specifically, the intrinsic distribution of planet sizes measured
by Kepler peaks at $\approx 2.4\,R_\oplus$, with a valley at
1.8\,$R_\oplus$, and falling off past $\approx$3.5\,$R_\oplus$
\citep[][Figure~7]{Fulton_et_al_2017}.

Figure~\ref{fig:rp_period_age} shows the sizes, orbital periods, and ages of
known transiting planets, emphasizing those with precise ages.
Considering planets younger than $10^8$ years, the smallest previously
known were AU~Mic~c ($\approx3.0\pm0.2\,R_\oplus$, averaging the
results of \citealt{martioli_aumicbc_2021} and
\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
\citealt{plavchan_planet_2020}).
The other transiting planets younger than $10^8$ years are much
larger, with sizes between 4 and 10$\,R_\oplus$
\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020}.

@@

We have argued that Kepler-1643~b, KOI-7368.01, and KOI-7913 {\it i)}
are 40 to 50 million years old, and {\it ii)} that each signal
is most likely planetary.  The evidence for the planetary
interpretation comes from an application of \texttt{VESPA} to the
Kepler data, alongside new adaptive optics imaging from NIRC2.  The
validity of the \texttt{VESPA} framework rests on the premise that
non-astrophysical false positives, including false alarms and known
neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
KOI-7368 both pass all but one of the canonical vetting checks: the
check for centroid offsets between the in and out-of-transit phases.
For both cases, we have argued the weak observed shifts are most
likely caused by starspot-induced variability in specific quarters
shifting the stellar center-of-light.

Each system has multiple indicators of youth that support the
reported ages.  For Kepler-1643, perhaps the strongest youth indicator
is its physical and kinematic association with RSG-5.  Based on the
color-absolute magnitude diagram, we are able to select members
of this cluster with a false positive rate of $\ll 5\%$
(Figure~\ref{fig:age}); Kepler-1643 is one such member.  The stellar
rotation period period agrees with this assessment, and {\bf its
lithium is a little low}.

The spatio-kinematic argument for the youth of KOI-7368 and
KOI-7913 are weaker because they are in an association of stars, CH-2,
that is more diffuse.  For KOI-7913, stronger indicators of its age
perhaps come from its binary.  Both stellar components in KOI-7913
have isochronal ages consistent with $40$\,Myr, but the isochronal
measurement for the secondary is more constraining since it is less
massive.  Both components also show H$\alpha$ in emission, which for
the $\approx$K6V primary is a good indicator that the star is
$\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
massive, the Li 6708\AA\ EW measurement and stellar rotation period
are more constraining, and provide independent verification of the
star's youth.
